---
title: "GoTChA_pipeline"
output: html_document
---

# Running Gotcha with parallel computing without using rslurm

This is a tutorial for processing [GoT-ChA](https://www.biorxiv.org/content/10.1101/2022.05.11.491515v1.full.pdf) genotyping libraries. The Gotcha R package required for this tutorial can be downloaded [here](https://github.com/landau-lab/Gotcha) 

## Use of BatchMutationCalling function

### First, we load the Gotcha R library:

```{r}
library(Gotcha)
```

### We are now ready to run the pipeline

```{r}
# Split the fastq files in chunks to allow for parallel processing and filter out reads that contain low quality scores,
# particularly at the mutation site of interest

path <- "/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/Gotcha_test_pipeline/GoTChA_test_data/"

FastqFiltering(out = path, min.quality = 15, which.read = "R1", ncores = 1, read.region = 15, min.bases= 5)

FastqSplit(path = path, out = path, reads = 5e+05, ncores = 12)

# Run Gotcha with parallel computing: submit one function call (1 core) per fastq chunk
BatchMutationCalling(out  = path,
                     which.read = "R1",
                     mutation.start = 1, # take the whole read
                     mutation.end = 50,
                     barcodes.file.path = "/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/Gotcha_test_pipeline/CBNN_0_singlecell.csv",
                     wt.sequence = "GCATGTATGCAATGCCTTGGTAGGAATGGGACAGGTGTAGGATGGAAAAT",
                     mut.sequence= "AAGGCGTTTCTTCTCTGACCGCACAACTGGGGCCTGGGGGGCTCCAAAGC",
                     wt.max.mismatch = 0, # set up on perfect matching for increased accuracy
                     mut.max.mismatch = 0,
                     ncores = 44,
                     rewrite = FALSE,
                     testing = TRUE,
                     max.distance = 2 
)

```

### Once the submitted jobs are completed, we can merge the outputs into one single data frame using the MergeMutationCalling function. 
```{r}
# Merge output from each BatchMutationCalling job. 
# Generate a new folder containing a .Rdata class object that can be directly loaded into R
MergeMutationCalling(out = path)
```

### We can then load the output from the MutationCalling function
```{r}
out_collapsed_tet2 <- readRDS("/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/Gotcha_test_pipeline/GoTChA_test_data/Split/Filtered/MergedOuts/outs.collapsed.Rdata")
out_collapsed_tet2$Sample <- "FRIZ01"
```

### Filter to barcodes of interest for genotyping - those used in the ATAC seurat object
### This is not strictly necessary, you could also genotype all of your cells
```{r}
FRIZ01_seurat_obj <- LoadH5Seurat("/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/Gotcha_test_pipeline/FRIZ01_CD56C_filt.h5Seurat")
cell_barcodes_filt <- Cells(FRIZ01_seurat_obj)
# subset to those barcodes
out_collapsed_tet2_filt  <- out_collapsed_tet2[which(out_collapsed_tet2$matched_barcodes %in% cell_barcodes_filt),]
# set column names for input to genotyping function
colnames(out_collapsed_tet2_filt) <- c("WhiteListMatch", "WTcount", "MUTcount", "MUTfraction", "Sample")
write.csv(out_collapsed_tet2_filt, "/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/out_collapsed_tet2_filt.csv")
```


### And run the GotchaLabeling function for noise correction and genotype labeling

## Noise Correction and Genotype Labeling
```{r}
#Use wild type and mutant read counts to predict genotype
genotypes_tet2 <- GotchaLabeling(path = "/sc/arion/projects/MDS/noelle/gotcha_pipeline_nwheeler/", infile = "out_collapsed_tet2_filt.csv", gene_id = "TET2", sample_id = "FRIZ01_TET2") # output is in 'path'/FRIZ01_TET2
```

### A folder named 'FRIZ01_TET2' (sample_id) is created in the path containing the output from genotyping!!
### Genotype calls will be in file 'FRIZ01_TET2/FRIZ01_TET2_genotype_labels.csv'


